---
title: "Palmer Penguins Statistical Analysis"
author: "Student"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Introduction
This report provides a comprehensive statistical analysis of the **Palmer Penguins** dataset. The objective is to investigate whether physical characteristics such as body mass and flipper length differ significantly across penguin species and sex.

### Dataset Overview
The dataset contains observations of penguins from three species (Adélie, Chinstrap, Gentoo) collected in the Palmer Archipelago, Antarctica.

```{r load_libraries}
library(palmerpenguins)
library(dplyr)
library(ggplot2)
library(e1071)
library(effsize)
library(car)

# Load dataset and remove NA values
data("penguins")
df <- na.omit(penguins)
```

---

## 1. Descriptive Statistical Analysis
We perform a descriptive analysis of the `body_mass_g` variable.

```{r desc_stats}
calc_stats <- function(x) {
  c(
    Mean = mean(x), Median = median(x),
    Min = min(x), Max = max(x),
    Var = var(x), SD = sd(x),
    Q1 = quantile(x, 0.25), Q3 = quantile(x, 0.75),
    IQR = IQR(x), Skewness = skewness(x), Kurtosis = kurtosis(x)
  )
}

# Overall Statistics
overall_stats <- calc_stats(df$body_mass_g)
knitr::kable(as.data.frame(overall_stats), col.names = "Overall Body Mass (g)")

# Species-wise Statistics
species_stats <- df %>%
  group_by(species) %>%
  summarise(
    Mean = mean(body_mass_g),
    Median = median(body_mass_g),
    Min = min(body_mass_g),
    Max = max(body_mass_g),
    Var = var(body_mass_g),
    SD = sd(body_mass_g),
    Q1 = quantile(body_mass_g, 0.25),
    Q3 = quantile(body_mass_g, 0.75),
    IQR = IQR(body_mass_g),
    Skewness = skewness(body_mass_g),
    Kurtosis = kurtosis(body_mass_g)
  )
knitr::kable(species_stats, digits = 2)
```

### Visualizations

```{r visuals_1, fig.width=10, fig.height=4}
p1 <- ggplot(df, aes(x = body_mass_g)) +
  geom_histogram(fill = "skyblue", color = "black", bins = 30) +
  theme_minimal() + labs(title = "Histogram of Body Mass")

p2 <- ggplot(df, aes(x = species, y = body_mass_g, fill = species)) +
  geom_boxplot() + theme_minimal() + labs(title = "Boxplot by Species")

p3 <- ggplot(df, aes(x = body_mass_g, fill = species)) +
  geom_density(alpha = 0.5) + theme_minimal() + labs(title = "Density Plot by Species")

cowplot::plot_grid(p1, p2, p3, ncol = 3)
```

---

## 2. Hypothesis Testing
Investigating whether male and female penguins have significantly different mean body masses.

**Null Hypothesis ($H_0$):** Mean body mass of males = Mean body mass of females.
**Alternative Hypothesis ($H_1$):** Mean body mass of males $\neq$ Mean body mass of females.

### Normality Check
```{r norm_check}
male_mass <- df$body_mass_g[df$sex == "male"]
female_mass <- df$body_mass_g[df$sex == "female"]

shapiro.test(male_mass)
shapiro.test(female_mass)
```
*Interpretation: The Shapiro-Wilk test indicates whether the data deviates significantly from normality.*

```{r qq_plots, fig.width=8, fig.height=4}
par(mfrow=c(1,2), mar=c(4,4,2,1))
qqnorm(male_mass, main = "QQ Plot - Male Body Mass")
qqline(male_mass, col = "red")
qqnorm(female_mass, main = "QQ Plot - Female Body Mass")
qqline(female_mass, col = "red")
par(mfrow=c(1,1), mar=c(5,4,4,2) + 0.1)
```

### Independent Two-Sample t-test
```{r ttest}
t_test_result <- t.test(body_mass_g ~ sex, data = df)
t_test_result

# Effect Size
cohen.d(body_mass_g ~ sex, data = df)
```
*Interpretation: A p-value < 0.05 indicates a significant difference between male and female body mass. Cohen's d provides the magnitude of this difference.*

---

## 3. One-Way ANOVA
Investigating whether mean body mass differs among the three penguin species.

### Assumptions Check
```{r assumptions_anova}
leveneTest(body_mass_g ~ species, data = df)
```

### ANOVA
```{r anova}
anova_res <- aov(body_mass_g ~ species, data = df)
summary(anova_res)
```

### Post-hoc Analysis
```{r posthoc}
tukey_res <- TukeyHSD(anova_res)
tukey_res
plot(tukey_res)
```

---

## 4. Non-Parametric Analysis
If assumptions of ANOVA are violated, we use the Kruskal-Wallis test.

```{r kruskal}
kruskal.test(body_mass_g ~ species, data = df)
```
*Interpretation: Comparing this p-value to the one-way ANOVA helps confirm the robustness of our conclusions.*

---

## 5. Two-Way ANOVA
Investigating the simultaneous effect of species and sex on penguin body mass.

```{r twoway_anova}
twoway_anova <- aov(body_mass_g ~ species * sex, data = df)
summary(twoway_anova)
```

### Sex-wise Boxplot
```{r boxplot_sex}
ggplot(df, aes(x = species, y = body_mass_g, fill = sex)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "Boxplot of Body Mass by Species and Sex", x = "Species", y = "Body Mass (g)")
```

---

## 6. Additional Analysis (Flipper Length)
Repeat the ANOVA analysis using `flipper_length_mm`.

```{r flipper_anova}
anova_flipper <- aov(flipper_length_mm ~ species, data = df)
summary(anova_flipper)

tukey_flipper <- TukeyHSD(anova_flipper)
tukey_flipper

ggplot(df, aes(x = species, y = flipper_length_mm, fill = species)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "Boxplot of Flipper Length by Species", x = "Species", y = "Flipper Length (mm)")
```
*Interpretation: The findings indicate whether flipper length shows consistent variations across species compared to body mass.*

---
## Conclusion
In this analysis, descriptive statistics highlighted differences in body mass between penguin species and sexes. Hypothesis testing confirmed significant differences between males and females. ANOVA tests (One-way and Two-way) combined with Tukey's HSD post-hoc tests further validated that species have significantly distinct body masses and flipper lengths.
